Controlling the dewetting morphologies of thin liquid films by switchable substrates

### 2022-12-09 Reply_RefereeA.tex

\section*{Reply to Referee A}
We thank the Referee for his/her report on our manuscript. 
We are glad to read that he/she finds it of potential interest and in principle 
suitable for publication, if a thorough revision is performed. 
We substantially improved and revised the paper for the resubmission and hope that we were able to clarify the obscure points, following all the Referee's remarks.
Hereafter, we provide answers to the various questions/comments raised. All changes in the revised version are reported in red color.

\begin{itemize}

\item[ \textbf{\underline{Comment 1.}}]
{ 
I am unable to follow most of the arguments made with respect to the disjoining pressure. 
For me, one needs to integrate the disjoining pressure to arrive at an effective interface potential, and 
I do not really know how to go from the effective interface potential to the surface tension. 
The equation (3) therefore merits much more discussion.
}

\item[ \textbf{{Answer}}]
{
We did not give, in fact, any particular argument with respect to the disjoining pressure, since 
we assumed that it is a very well established concept employed in any theoretical and computational study of thin liquid film hydrodynamics in the lubrication approximation framework.
We also assumed that an equation like Eq.(3) (now Eq.(2)), which is approximately twenty-five years old \cite{SCHWARTZ1998173,Mitlin94}, did not deserve a lengthy discussion.
However, our assumptions were probably erroneous and, therefore, we give, here, some more explanation. 
Let us, first, make a step back and underline a fundamental difference between the physics of {\it one free interface} and that of a {\it film}:
the latter is a layer of liquid confined between {\it two} interfaces. When the two interfaces are brought close enough to each other (the layer is thin), they interact.
The disjoining pressure stems from this interaction. Thermodynamically, it represents minus the derivative of the free energy difference, per unit surface, between 
the film and the bulk of the same liquid, or, equivalently, the pressure difference inside and outside the film 
\cite{Deryaguin1940,DeryaguinChuraev1978}.
In the case of wetting, there can be three types of interfaces: liquid/air, liquid/solid (the substrate) and, 
if the substrate is only partially wetted, solid/air, thus giving rise to three-phase contact lines. 
The usually called {\it surface tension} is the energy cost of forming a liquid/air interface of unit area and its only knowledge is, therefore, not enough in our case. 
The Young's condition \cite{Young1805,DeGennes1985} connects the surface tension with the analogous energies per unit
area associated with the solid/liquid ($\gamma_{\text{sl}}$) and solid/gas ($\gamma_{\text{sg}}$) interface as
\begin{equation}
\cos \theta = \frac{\gamma_{\text{sg}}-\gamma_{\text{sl}}}{\gamma},
\end{equation}
where $\theta$ is the contact angle.
So, there is no need (and no way) to go from the disjoining pressure (or the interface potential) to the surface tension, since the two quantities carry complementary physical information,
and both are required to determine the dynamics of dewetting.
Of course, on the other hand, the disjoining pressure depends also on the surface tension, which appears, in fact, in its expression (2). 
We think, though, that part of the confusion might be due to the fact that we do not define the symbol of the surface tension at its first occurrence, but only later. 
We have now fixed this and we apologize for it.
}

@@

\item[ \textbf{\underline{Comment 4.}}]
{ 
In addition, there is a lot of old literature that relates the dewetting to the disjoining pressure, and shows that the nucleation of a hole in a wetting film is much more difficult than the nucleation of such a wetting film; there are experiments on this by Rutledge and Taborek, theory by Schick and Taborek and Bausch, Blossey and Indekeu.
}

\item[ \textbf{{Answer}}]
{
We are well aware of the literature quoted by the Referee, but these are works 
about {\it prewetting}, namely the phenomenon of {\it nucleation} (condensation)
of films/droplets on solid surfaces confining a critical liquid-vapour mixture
{\it as a function of temperature}. So, we are a bit puzzled about what such literature may have to do with the {\it dewetting} of a film on a chemically patterned substrate. Here, in fact, we are modelling a thin liquid film which is 
thermodaynamically stable against evaporation (i.e. one can consider that 
we are at a - fixed - temperature below the critical one).
}

\item[ \textbf{\underline{Comment 5.}}]
{ 
The explanations also lack a discussion on time, length, and energy scales. 
I have no clue what the values mean ``the numerical values, in LB units, are set to $\gamma= 0.01$ and $\mu = 1/6$''?. 
}

\item[ \textbf{{Answer}}]
{
Length and time scales, as discussed in Comment 2 and as mentioned also in the original version of the manuscript, 
are made dimensionless with the mean film height and the characteristic time $t_0$. 
These two quantities 
have a clear physical meaning and can, therefore, be used to compare with 
experimental setups. In particular, the expression of $t_0$, through its dependence 
on $q_0$ (the wavenumber of the most unstable mode) contains the details 
of the disjoining pressure, such as the Hamaker constant $A_H$. For instance, let us consider the problem studied in \cite{becker2003complex}, the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 
dioxide substrate. The Hamaker constant for this system is 
$A_H  = 2.2 \times 10^{-20} \, \text{J}$ and the surface tension is 
$\gamma = 0.03 \, \text{N}/\text{m}$, such that, 
from equation (3) of \cite{becker2003complex} and 
from the definition of $q_0^2 = \frac{1}{2\gamma} \Pi^{\prime}(h_0)$, 
we get (neglecting the short range part of the interface potential)
$$
q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}.
$$
The characteristic time of the growth of the instability, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)
$$
t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}
$$
and it can be seen that indeed film rupture occurs within time 
scales of the order of $t_0$ \cite{becker2003complex}.
We have reported these comments on comparison with actual experimental 
values in the new version of the paper as follows:\\
\\
\textcolor{red}{To give a flavour of how to compare with actual experimental systems, let us notice
first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
 $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
 the Hamaker constant $A_H$. Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 
dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
The characteristic time of the growth of the instability for these experiments, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 
$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$) $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ and it can be seen that indeed film rupture occurs within time 
scales of the order of $t_0$ \cite{becker2003complex}, as in our 
simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
with homogeneous substrates~\cite{PhysRevE.104.034801}).} 
\\

### 2022-12-09 Reply_RefereeB.tex

\item[ \textbf{\underline{Comment 3.}}]
{
The relevance of the manuscript relies on the significance of the substrate's switchable properties. 
Particularly, it is assumed that the wetting properties can be modulated on space but also on time. 
The authors have mentioned in the introduction that this can be potentially done. 
However, it would be desirable that the authors relate their model of switchable substrate with some specific experimental realization, where the actual length and time scales are mentioned.
}

\item[ \textbf{{Answer}}]
{
%This is indeed a very interesting comment.
%Although our approach, as theorists/numericists, can be seen more of {\it theoretical
%engineering} and our picture would be, then, unlikely devoid of flaws, we are glad 
%to try to envisage a potential experiment.
%\JH{??}
%We thank the Referee for bringing this up.
%First, we like to note that experiments are not our field of expertise and therefore %our envisioned realization could be flawed.
%The two relevant scales are $t_0$ and $q_0$ (or $2\pi/q_0$).
%Considering the experimental parameters from~\cite{becker2003complex,PhysRevLett.99.114503} %we get, 
%\begin{align*}
%    t_0 &\approx 300\,\text{s},\qquad~~~\qquad \frac{2\pi}{q_0}\approx 
%    400\,\text{nm}.
%\end{align*}
%The time scale is in the region of minutes and the wettability pattern wavelengths are %in the interval $(800\,\text{nm}, 4000\,\text{nm})$. 
A possibility could be to prepare a thin liquid film on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
external stimuli (a light emitter).

\textcolor{red}{An ideal candidate could be a digital multimirror device (DMD).
This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
of polystyrene deposited on an oxidized silicon wafer,  
we evaluate the retraction speed 
$U_{\Theta} = \frac{\Theta^3 \gamma}{9 \mu}$ to be 
$U_{\Theta} \approx 10^{-2} \, \mu \text{m}/\text{s}$. 
The (minimum) {\it pattern speed} can be 
estimated, from the pixel size with frame rate $\sim 1 s^{-1}$, as 
$v_{\theta} \sim 10 \, \mu \text{m}/\text{s}$, that would give $\Gamma \sim 10^3$, i.e. well within 
the rivulet regime ($\Gamma > 1$). We note, also, that both $U_{\Theta}$ and $v_{\theta}$ can 
be widely modulated: the former, playing either with the (temperature and molecular weight dependent) viscosity 
or tailoring the substrate to making it more hydrophobic (i.e. increasing the contact angle); the latter, tuning
the spatial resolution and frame rate of the DMD. We expect, thus, the range of achievable $\Gamma$'s to 
be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values.}

We have included these comments in the revised manuscript just before the conclusions.

### 2022-12-09 letter_to_editor.tex
Resubmission, waiting for Jens

\noindent Dear Editor,\\
\\
\noindent We gratefully thank you for sending us the reports on our manuscript. We are glad to see 
that the Referees agree that the work is interesting and scientifically relevant, with one of them 
(Referee B) inclined to accept it. 
\textcolor{red}{In particular, we would like to underline that not a single 
concern was raised on the physics, namely on the reliability of the results.
On the other hand, we acknowledge that certain points were not explained and/or discussed adequately, so we have put a lot of effort in clarifying all
doubts and concerns.}
We have made, therefore, a major revision taking into account all Referees' reasonable remarks, including novel 
results from new simulations and theoretical derivations (see the new Supplementary Material).
Let us, please, stress further how and why we deem that the present paper meets the Physical
Review Letters criteria of impact and innovation (interest, as we already noted, was acknowledged by the Referees).
We have found a novel effect in the dewetting on substrates with properly tailored time dependent properties, and 
suggested how to exploit it to control the dewetting morphologies, 
\textcolor{red}{even proposing, motivated by one Referee's suggestion, a possible experimental realisation.}
We think that the impact of this discovery is out of question, as witnessed also by the various publications, also on in PRL on dewetting and closely related problems
(just to cite but a few recent ones: J.C. Fern\'andez-Toledano et al, ``How wettability controls nanoprinting'',
{\it Phys. Rev. Lett.} {\bf 124} 224503 (2020); A.A. Pahlavan et al,
``Evaporation of binary-mixture liquid droplets: the formation of picoliter pancakelike shapes'', {\it Phys. Rev. Lett.} {\bf 127}, 024501 (2021); 
C. Clavaud et al, ``Modification of the fluctuation dynamics of ultrathin wetting films'', {\it Phys. Rev. Lett.} {\bf 126}, 228004 (2021)).
What it is somehow questionable, instead, is the trustworthiness of some of the Referees in judging a paper on microscale multiphase fluid dynamics: one of 
them (Referee A) ignores the concept of disjoining pressure, confuses dewetting with prewetting and concepts from the physics of free surface flows to wetting.
The main ``argument'' of Referee C against publication is that he/she does not see ``how would the transport of a fluid by rivulets 
surpass the transport by droplets''; but the paper is about controlling morphologies, not about transport.\\
\textcolor{red}{We believe, then, that our manuscript, in its new version, deserves to be brought again to your attention as a contribution to Physical Review Letters, 
confident that it might be now suitable of publication.}\\
\\
With our best regards,\\
the Authors.

>>>>

\noindent Dear Editor,\\
\\
\noindent We gratefully thank you for sending us the reports on our manuscript. We are glad to see 
that the Referees agree that the work is interesting and scientifically relevant, with one of them 
(Referee B) inclined to accept it. 
\textcolor{red}{In particular, we would like to underline that not a single 
concern was raised on the physics, namely on the reliability of the results.
On the other hand, we acknowledge that certain points were not explained and/or discussed adequately, so we have put a lot of effort in clarifying all
doubts and concerns.}
We have made, therefore, a major revision taking into account all Referees' reasonable remarks, including novel 
results from new simulations and theoretical derivations (see the new Supplementary Material).
Let us, please, stress further how and why we deem that the present paper meets the Physical
Review Letters criteria of impact and innovation (interest, as we already noted, was acknowledged by the Referees).
We have found a novel effect in the dewetting on substrates with properly tailored time dependent properties, and 
suggested how to exploit it to control the dewetting morphologies, 
\textcolor{red}{even proposing, motivated by one Referee's suggestion, a possible experimental realisation.}
We think that the impact of this discovery is out of question, as witnessed also by the various publications, also in PRL on dewetting and closely related problems
(just to cite but a few recent ones: J.C. Fern\'andez-Toledano et al, ``How wettability controls nanoprinting'',
{\it Phys. Rev. Lett.} {\bf 124} 224503 (2020); A.A. Pahlavan et al,
``Evaporation of binary-mixture liquid droplets: the formation of picoliter pancakelike shapes'', {\it Phys. Rev. Lett.} {\bf 127}, 024501 (2021); 
C. Clavaud et al, ``Modification of the fluctuation dynamics of ultrathin wetting films'', {\it Phys. Rev. Lett.} {\bf 126}, 228004 (2021)).
What it is somehow questionable, instead, is the trustworthiness of some of the Referees in judging a paper on microscale multiphase fluid dynamics: one of 
them (Referee A) ignores the concept of disjoining pressure, confuses dewetting with prewetting and concepts from the physics of free surface flows to wetting.
The main ``argument'' of Referee C against publication is that he/she does not see ``how would the transport of a fluid by rivulets 
surpass the transport by droplets''; but the paper is about controlling morphologies, not about transport.\\
\textcolor{red}{We believe, then, that our manuscript, in its new version, deserves to be brought again to your attention as a contribution to Physical Review Letters, 
confident that it might be now suitable of publication.}\\
\\
With our best regards,\\
the Authors.

### 2022-12-09 main_AS_final.tex
Resubmission, waiting for Jens

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\definecolor{jlblue}{rgb}{0.0,0.6056031611752245,0.9786801175696073}
\definecolor{jlorange}{rgb}{0.8888735002725198,0.43564919034818994,0.2781229361419438}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

@@

\textcolor{red}{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 
The external stimulus, our moving pattern, can be realized, e.g., by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}

@@

scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}, as in our 

@@

    with $v_{\theta}= 0$ and $\lambda= 512 h_0$ (blue circles), $\lambda=256 h_0$ (orange squares) and $\lambda=170 h_0$ (green stars).
    The dashed lines indicate the geometrically expected values of the droplet height $h_d$, assuming monodispersity and a perfect spherical cap shape. 

@@

For $v_{\theta} = 0.2 v_0$ we still observe the formation of droplets, similarly to the stationary case ($v_{\theta} = 0$). However, these are transported with the contact angle minima, reproducing a somehow similar behaviour recently described in a numerical study of a droplet on a moving wettability step~\cite{GrawitterStark1}.
If the pattern speed is further increased, for $v_{\theta} = 2 v_0$ we observe the development of rivulet-like structures, aligned with $\mathbf{v}_{\theta}$. 

@@

\noindent {\it Acknowledgements}. SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11 and Project-ID 431791331—CRC1452 ``Catalysis at Liquid Interfaces''.

>>>>

\definecolor{pyblue}{HTML}{1F77B4}
\definecolor{pyorange}{HTML}{FF7F0C}
\definecolor{pygreen}{HTML}{2CA02C}
\definecolor{pyred}{HTML}{D62728}
\definecolor{jlblue}{rgb}{0.0,0.6056031611752245,0.9786801175696073}
\definecolor{jlorange}{rgb}{0.8888735002725198,0.43564919034818994,0.2781229361419438}
\definecolor{jlgreen}{rgb}{0.2422242978521988,0.6432750931576305,0.3044486515341153}
\definecolor{jlviolet}{rgb}{0.7644401754934356,0.4441117794687767,0.8242975359232758}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

@@

% \textcolor{red}{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 
% The external stimulus, our moving pattern, can be realized, e.g., by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}

@@

scales of the order of $t_0$~\cite{becker2003complex}, as in our 

@@

    with $v_{\theta}= 0$ and $\lambda= 512 h_0$ (\textcolor{jlblue}{$\bullet$}), $\lambda=256 h_0$ (\textcolor{jlorange}{$\blacksquare$}) and $\lambda=170 h_0$ (\textcolor{jlgreen}{$\star$}).
    % The dashed lines indicate the geometrically expected values of the droplet height $h_d$, assuming monodispersity and a perfect spherical cap shape. 

@@

For $v_{\theta} = 1.7 v_0$ we still observe the formation of droplets, similarly to the stationary case ($v_{\theta} = 0$). However, these are transported with the contact angle minima, reproducing a somehow similar behaviour recently described in a numerical study of a droplet on a moving wettability step~\cite{GrawitterStark1}.
If the pattern speed is further increased, for $v_{\theta} = 17.5 v_0$ we observe the development of rivulet-like structures, aligned with $\mathbf{v}_{\theta}$. 

@@

\noindent {\it Acknowledgements}. SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11 and Project-ID 431791331—CRC1452 ``Catalysis at Liquid Interfaces''. SZ acknowledge financial support from the Independent Research Fund Denmark through a DFF Sapere Aude Research Leader grant (grant number 9063-00018B).
